A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-07-26 text/discussion.Rmd
update Breast histological subtypes discussion, add review references

We find that most tumors segregate into one of twelve biologically distinct
clusters, regardless of the cancer type (**figure 1G,I** and **supplementary
figure S1**). Some cancer types ( _e.g._ pancreatic adenocarcinomas in the RAS
arm clusters 8 and 9, and Kidney Chromophobe in cluster 1, see **supplementary
figure S1**) are unevenly distributed, though we could not identify cancer type
that was exclusive to a single tumor class. Only cancer types with the fewest
samples were found to be absent from one or more classes at all. Somewhat
surprisingly to us, this finding extends to histological subtypes of breast,
head and neck cancers, leukemias, _etc._ (see **Figure 1J** and **Supplemental
Figure S2**). This result implies that many, if not most, histological subtypes
primarily reflect differences in cell-of-origin, rather than fundamental
differences in proximal cause. It has been hypothesized that differences in the
molecular regulators of development in the array of cell types present in the
breast epithelium drive the various histological phenotypes [@breast-subtypes].
Consistent with this view we found that breast tumor samples of the Luminal A
subtype were heavily biased toward membership in clusters 1 and 7, and basal
tumor samples were biased toward cluster 6 (**figure 1J**). However, both
subtypes contained samples in every other class without exception, and Luminal B
and Her2 positive samples are distributed across clusters. We interpret these
data to mean that cell-of-origin molecular machinery can predispose certain cell
types within the breast epithelium to forming tumors of one class or another but
is not determinative; this view is compatible with the previously stated
hypothesis but opens the way for a more granular view of individual tumors.
Indeed, it would be surprising if we did not observe bias for some cancers and
subtypes amongst our classes, since some treatment regimens have greater
efficacy for a significant fraction of patients of a given cancer/histological
subtype. Nonetheless, the basis for some tumors being treatment-refractory in
spite of receiving the standard of clinical care for diagnostic markers remains
elusive. Doubtless some of this is due to bad luck, as tumors can metastasize
and remain dormant years before they are detected at distal sites, or resistant
clones may have already arisen at undetectable levels [@hallmarks_II], but our
analysis suggests the possibility of identifying more informative molecular,
histological or cellular subtypes that could form a basis for future separation
of patients into different precision treatment regimens.

>>>>

We find that most tumors segregate into one of twelve biologically distinct
clusters, regardless of the cancer type (**figure 1G,I** and **supplementary
figure S1**). Some cancer types ( _e.g._ pancreatic adenocarcinomas in the RAS
arm clusters 8 and 9, and Kidney Chromophobe in cluster 1, see **supplementary
figure S1**) are unevenly distributed, though we could not identify cancer type
that was exclusive to a single tumor class. Only cancer types with the fewest
samples were found to be absent from one or more classes at all. Somewhat
surprisingly to us, this finding extends to histological subtypes of breast,
head and neck cancers, leukemias, _etc._ (see **Figure 1J** and **Supplemental
Figure S2**). This result implies that many, if not most, histological subtypes
could reflect differences in cell-of-origin, rather than fundamental differences
in proximal cause. The five major subtypes of breast cancer have histological
and molecular correlates that define them and how they respond to experimental
stimuli [@breast-subtypes2]. It has been hypothesized that differences in the
molecular regulators of development in the precursor cell types present in the
breast epithelium drive the various histological phenotypes [@breast-subtypes;
@breast-subtypes3]. Consistent with this view we found that breast tumor samples
of the Luminal A subtype were heavily biased toward membership in clusters 1 and
7, and basal tumor samples were biased toward cluster 6 (**figure 1J**).
However, both subtypes also contained samples in every other class (without
exception), and Luminal B and Her2 positive samples are distributed across
clusters. We interpret these data to mean that inherited cell-of-origin
molecular profiles may predispose certain precursor cell types within the breast
epithelium to forming tumors of one class or another but is not determinative;
this view is compatible with the previously stated hypothesis but opens the way
for a more granular view of individual tumors. It would be surprising if we did
not observe bias for some cancers and subtypes amongst our classes, since some
treatment regimens have greater efficacy for a significant fraction of patients
of a given cancer/histological subtype [@breast-subtypes2]. Nonetheless, the
basis for some tumors being treatment-refractory in spite of receiving the
standard of clinical care for diagnostic markers remains elusive. Doubtless some
of this is due to bad luck, as tumors can metastasize and remain dormant years
before they are detected at distal sites, or resistant clones may have already
arisen at undetectable levels [@hallmarks_II], but our analysis suggests the
possibility of identifying more informative molecular, histological or cellular
subtypes that could form a basis for future separation of patients into
different precision treatment regimens.
